Paper Analysis

### 2025-04-22 Paper_WriteUp/Paper_Plots.qmd
updating figs with revised models

#read in models
fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))

@@

ggsave(plot = comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients 17-19}
#| eval: false
gamma_ests <- model_17_19$summary(variables = 
                                              c("gamma1",
                                                "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests[1])))

@@

```{r interpret the estimated coefficients gm}
#| eval: false
gamma_ests_gm <- model_gm$summary(variables = c("gamma1",
                                                          "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests_gm[1])))
## change in win probability for elo difference
1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
```

@@

```{r figure 4 left}
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.1, 0.1)) +
  theme_single_y()

ggsave(plot = winner_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_winner.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(),
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.1, 0.1)) +
  theme_single_y()

ggsave(plot = winner_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_winner.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 right - plot player effects}
#| eval: false
## then repeat for player effects
indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

indiv_plot_gm

@@

#getting the draws
mu_beta_draws = mu_beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
  rename(draw = mu_beta) %>%
  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
         dataset_id = as.numeric(dataset_id),
         true_mu_beta = case_when(dataset_id %in% 1:3 ~ true_mu_betas[1],
                                  dataset_id %in% 4:6 ~ true_mu_betas[2],
                                  dataset_id %in% 7:9 ~ true_mu_betas[3]),
         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
                               dataset_id %in% c(2, 5, 8) ~ 5000,
                               dataset_id %in% c(3, 6, 9) ~ 20000),
         true_mu_beta = factor(true_mu_beta, levels = c(as.character(true_mu_betas))),
         num_games = factor(num_games, levels = c("1000", "5000", "20000")))

#for the geom_segment below
true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
  mutate(ymin = as.numeric(factor(true_val)) - 0.25,
         ymax = as.numeric(factor(true_val)) + 0.25)

#plotting
mu_beta_plot = mu_beta_draws %>% ggplot(aes(y = true_mu_beta, x = draw, colour = num_games)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  geom_segment(data = true_mu_betas_plotting,
               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.5) +
  labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
  theme_single_y_legend() +
  scale_color_viridis_d(option = "D")

>>>>

#read in models
fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_n1.RDS"))
fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_n1.RDS"))
fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_n1.RDS"))
fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_n1.RDS"))

@@

# ggsave(plot = comp_global, 
#        filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
#        height = 5, width = 7)
```

```{r interpret the estimated coefficients 17-19}
#| eval: false
gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
                                              c("gamma1",
                                                "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests[1])))

@@

```{r interpret the estimated coefficients gm}
#| eval: false
gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
                                                          "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests_gm[1])))
## change in win probability for elo difference
1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
```

@@

```{r figure 4 left}
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_pointinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.125, 0.125)) +
  theme_single_y()

ggsave(plot = winner_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "fig4_left.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_pointinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(),
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_pointinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  xlim(c(-0.125, 0.125)) +
  theme_single_y()

ggsave(plot = winner_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "fig4_right.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 right - plot player effects}
#| eval: false
## then repeat for player effects
indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_pointinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

@@

#getting the draws
mu_beta_draws = mu_beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
  rename(draw = mu_beta) %>%
  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
         dataset_id = as.numeric(dataset_id),
         true_mu_beta = case_when(dataset_id %in% 1:3 ~ true_mu_betas[1],
                                  dataset_id %in% 4:6 ~ true_mu_betas[2],
                                  dataset_id %in% 7:9 ~ true_mu_betas[3]),
         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
                               dataset_id %in% c(2, 5, 8) ~ 5000,
                               dataset_id %in% c(3, 6, 9) ~ 20000),
         true_mu_beta = factor(signif(true_mu_beta, 3), levels = c(as.character(signif(true_mu_betas, 3)))),
         num_games = factor(num_games, levels = c("1000", "5000", "20000")))

#for the geom_segment below
true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
  mutate(ymin = as.numeric(as.factor(true_val)) - 0.3,
         ymax = as.numeric(as.factor(true_val)) + 0.3)

#plotting
mu_beta_plot = mu_beta_draws %>% ggplot(aes(y = true_mu_beta, x = draw, colour = num_games)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  geom_segment(data = true_mu_betas_plotting,
               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
               color = "red", linewidth = 1.25) +
  labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
  theme_single_y_legend() +
  scale_x_continuous(breaks = c(signif(true_mu_betas, 3)), 
                     labels = c(as.character(signif(true_mu_betas, 3)))) +
  scale_color_viridis_d(option = "D")
mu_beta_plot
